Syntactic realization with data-driven neural tree grammars

### 2016-10-11 paper/draft.tex
figure->Figure
spell check and typos

\blfootnote{
    % % final paper: en-us version (to licence, a license)
    %
    \hspace{-0.65cm}  % space normally used by the marker
    This work is licenced under a Creative Commons 
    Attribution 4.0 International License.
    License details:
    \url{http://creativecommons.org/licenses/by/4.0/}
}

@@

In lexicalized tree grammars, the lexicon and the grammatical rules
are one and the same.
%
The set of possible grammatical moves which can be made are simultaneously the set of possible words which can be used next.
%
This means inducing a tree grammar from a data set is a matter of inferring the set of constructions in the data.

@@

The simplest way to embed supertags is to treat each structure
as a distinct token that indexes a corresponding learned vector.  This
places no constraints on the learned similarity function, but it also
ignores the hierarchical structure of the elementary trees
themselves.  Previous work on deep learning with graph structures
suggests convolutional neural networks can exploit similarities in structure \cite{niepert2016learning,kalchbrenner2014convolutional}.  Thus, we developed
analogous techniques to encode supertags based on their underlying
tree structure.  In particular, to embed a supertag, we embed each
node, group the resulting vectors to form a tensor, and then summarize
the tensor into a single vector using a series of convolutional neural
networks.

@@

Our models predict supertags as a function of the target word and its
context.  Neural networks make it possible to generalize over such
contexts by learning to represent them with a hidden state vector that
aggregates and clusters information from the relevant history.  Our
approach is to do this using a recurrent tree network.
%
While recurrent neural networks normally use the previous hidden state in the sequential order of the inputs,
%require  the hidden state to be from the
%temporally previous hidden state,
recurrent tree networks use the hidden state from the parent.
%
Utilizing the parent's hidden state rather than the sequentially previous hidden
state, the recurrent connection can travel down the branches of a tree.
%
An example of a recurrent tree network is shown in figure \ref{fig:rtn}.

@@

The model is implemented as a feed-forward neural network.
%
Equation \ref{eq:fergusn} details the model formulation.
%
The lexical material, $lex_p$ and $lex_c$, are embedded using the word embedding matrix, $G_w$, concatenated, and mapped to a new vector, $\omega_{lex}$, with a fully connected layer, $FC_1$.
%
The child supertag, $tag_c$, is embedded with $G_\Lambda$ and concatenated the lexical vector, $\omega_{lex}$, forming an intermediate vector representation of the node, $\omega_{node}$.
%
The node vector is repeated for each of the parent's possible supertags, $tagset_p$, and then concatenated with their embeddings to construct the set of treelet vectors, $\Omega_{treelet}$.
%
The vector states for the leaf nodes are similarly constructed, but instead combine the lexical vector, $\omega_{lex}$ with the embeddings of the child's possible supertags, $tagset_c$.
%
The final operation induces a probability distribution over the treelet and leaf vectors using a score computed by the vectorized function, $\Psi_{predict}$, as the scalar in a softmax distribution. 
%
%\vspace*{-0.55cm}
\begin{align}
&\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
&\omega_{node}=concat([G_\Lambda(tag_c);~\omega_{lex}]) \nonumber \\
&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_\Lambda(tagset_p)]) \nonumber \\
&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_\Lambda(tagset_c)]) \nonumber \\
&P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
\frac{exp(\Psi_{predict}(\omega_{treelet_i})))}
{\sum_{j \in |tagset_p|} exp(\Psi_{predict}(\omega_{treelet_j})))} \nonumber \\
&P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
\frac{exp(\Psi_{predict}(\omega_{leaf_i})))}
{\sum_{j \in |tagset_c|} exp(\Psi_{predict}(\omega_{leaf_j})))} \nonumber 
\end{align}
%\vspace*{-0.45cm}

At generation time, we are given a full dependency tree.  A decoding
step is necessary to compute a high probability assignment for all
supertags simultaneously.
%
In this process, tags for children must be chosen consistently with
one another, and the resulting probabilistic information must be
propagated upward to rerank tags higher in the tree.
%
We solve this problem with an A* algorithm.
%
At each step, the algorithm uses a priority queue to select subtrees based on their inside-outside scores.
%
The inside score is computed as the sum of the log probabilities of the supertags in the subtree.
%
The outside score is the sum of the best supertag for nodes outside the subtree, similar to \newcite{lewis2014improved}.
%
Once selected, the subtree is attached to the possible supertags of its parent that are both locally consistent and consistent among its already attached children.
%
These resulting subtrees are placed into the priority queue and the algorithm iterates to progress the search.
%
The search succeeds when the first complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed.}

@@

The Fergus-R model uses only lexical information as input to calculate the probability distribution over each node's supertags. 
%
The specific formulation is detailed in Eq. \ref{eq:fergusr}.
%
First, a parent node's children, $lex_C$, are embedded using the word embedding matrix, $G_w$, and then summarized with an attention function, $\Psi_{attn}$, to form the child context vector, $\omega_{C}$. 
%
The child context is concatenated with the embedded lexical information of the parent node, $lex_p$, and mapped to a new vector space with a fully connected layer, $FC_1$, to form the lexical context vector, $\omega_{lex}$.
%
The context vector and a topological vector for indexing the internal tree state (see \ref{subsec:rtn}) are passed to the recurrent tree network, $RTN$, to compute the full state vector for the parent node, $\omega_{node}$.
%
Similar to Fergus-N, the state vector is repeated and concatenated with the vectors of the parent node's possible supertags, ${tagset}_p$, and mapped to a new vector space with a fully connected layer, $FC_2$.
%
A vector in this vector space is labeled $\omega_{elementary}$ because the combination of supertag and lexical item constitutes an elementary tree.
%
The last step is to compute the probability of each supertag using the vectorized function, $\Psi_{predict}$.
%
%\vspace*{-0.55cm}
\begin{align}
&\omega_{C} = \Psi_{attn}(G_w(lex_C)) \label{eq:fergusr} \\
&\omega_{lex} = FC_1(concat(\omega_{C},~G_w(lex_p))) \nonumber \\
&\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_\Lambda(tagset_p))) \nonumber \\
&P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
\frac{exp(\Psi_{predict}(\omega_{elementary_i})))}
{\sum_{j \in |\Omega|} exp(\Psi_{predict}(\omega_{elementary_j))}} \nonumber
\end{align}
%\vspace*{-0.45cm}

@@

The final step to linearizing the output of Fergus-N and Fergus-R---a dependency tree annotated with supertags and partial attachment information---is a search over possible orderings with a language model. 
%
There are many possible orderings due to adjunction operation
What remains to be determined is the order in which adjuncts are attached. 
%
Following \newcite{bangalore2000exploiting}, a language model is used to select between the alternate orderings. 
%
The language model used is a two-layer LSTM trained using the Keras library on
the surface form of the Penn Treebank.
%
The surface form was minimally cleaned\footnote{With respect to the surface form, the only
cleaning operations were to merge proper noun phrases into single tokens.  Punctuation and other
common cleaning operations were not performed.} to simulate realistic scenarios.

@@

Using the representations of \ref{sec:neural}, the models of
\ref{sec:models} can be instantiated in six different ways.  We can
use a feed-forward Fergus-N architecture or a recurrent Fergus-R
architecture.  Each architecture can embed supertags \emph{minimally},
by learning a scalar corresponding to each supertag; \emph{atomically}, by learning an
embedding vector corresponding to each supertag; or
\emph{structurally}, by using convolutional coding over each
supertag's tree structure to form a vector.
%
In each case, the vector (a size-one vector in the \emph{mimimal} condition) is concatenated as described in \ref{sec:models}.

@@

From this input, we extracted the atomic supertag prediction instances
and trained a network defined by each of the architectures of
\ref{sec:models} and each of the supertag representations of
\ref{sec:neural}.  As always, we used Section 02-21 for training,
Sections 22 for development, and Section 23 for testing.  A complete
description of network organization and training parameters is given
in the appendix.  The code and complete experimental setup are
publically
available.\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}

@@

Figure~\ref{table:linresults} shows the NLG evaluation results for the
different models.
%
All differences in model are significant using an Independent\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.} t-test ($p<10^{-5}$). 
%
For both models, the differences between structural embeddings (using convolutional coding) and atomic embeddings (using standard vector embedding techniques) were not significant, while the differences between the two embeddings and minimal embeddings were significant ($p<10^{-5}$). 
%
The performance confirms our expectation that differences in supertag
accuracy after decoding correlate with NLG accuracy overall, but that
differences in NLG performance are attenuated.
%
We note by comparison that Bangalore and Rambow report an accuracy of
74.9\% in their best evaluation of FERGUS---on a data set of just 100
sentences with an average length of 16.7.
%
Our evaluation, on 2400 sentences with an average length of 22, is
more strenuous.

@@

There are several lines of related work which explore stochastic tree
models from the standpoint of parsing and understanding.
%
While using the same methods, NLG has different goals and we think the perspective is instructive. 
%
Where parsing infers most probable underlying structure, generation infers the most likely way of expressing a semantic structure.
%
This divergence of goals leads to different concerns, alternatives, and emphasis. 
%---which unlabeled dependency trees have served as our proxy in this work.

The works most similar to ours explicitly model tree structures, but focus on resolving the uncertainty invovled with the latent structure of an observed sentence. 
%
For example, the top down tree structure of \newcite{zhang2016top} expresses the generation of a dependency structure as the decisions of a set of long short-term memory networks.  
%
For each decision, the possible options are different tree structures which can produce the target linear form.
%
In contrast, the generation problem is concerned with different linear forms that can result from the same tree structure. 
%
In more extensive tasks, the generation problem can include simulated interpretion to inform decisions; using the ease of structural inference from linear form quantifies the understandability of a sentence.

>>>>

\blfootnote{
    % % final paper: en-us version (to licence, a license)
    %
    \hspace{-0.65cm}  % space normally used by the marker
    This work is licensed under a Creative Commons 
    Attribution 4.0 International License.
    License details:
    \url{http://creativecommons.org/licenses/by/4.0/}
}

@@

In lexicalized tree grammars, the lexicon and the grammatical rules
are one and the same.
%
The set of possible grammatical moves which can be made are simultaneously the set of possible words which can be used next.
%
This means that inducing a tree grammar from a data set is a matter of inferring the set of constructions in the data.

@@

The simplest way to embed supertags is to treat each structure
as a distinct token that indexes a corresponding learned vector.  This
places no constraints on the learned similarity function, but it also
ignores the hierarchical structure of the elementary trees
themselves.  Previous work on deep learning with graph structures
suggests convolutional neural networks can exploit similarities in structure \cite{kalchbrenner2014convolutional,niepert2016learning}.  Thus, we developed
analogous techniques to encode supertags based on their underlying
tree structure.  In particular, to embed a supertag, we embed each
node, group the resulting vectors to form a tensor, and then summarize
the tensor into a single vector using a series of convolutional neural
networks.

@@

Our models predict supertags as a function of the target word and its
context.  Neural networks make it possible to generalize over such
contexts by learning to represent them with a hidden state vector that
aggregates and clusters information from the relevant history.  Our
approach is to do this using a recurrent tree network.
%
While recurrent neural networks normally use the previous hidden state in the sequential order of the inputs,
%require  the hidden state to be from the
%temporally previous hidden state,
recurrent tree networks use the hidden state from the parent.
%
Utilizing the parent's hidden state rather than the sequentially previous hidden
state, the recurrent connection can travel down the branches of a tree.
%
An example of a recurrent tree network is shown in Figure \ref{fig:rtn}.

@@

The model is implemented as a feed-forward neural network.
%
Equation \ref{eq:fergusn} details the model formulation.
%
The lexical material, $lex_p$ and $lex_c$, are embedded using the word embedding matrix, $G_w$, concatenated, and mapped to a new vector, $\omega_{lex}$, with a fully connected layer, $FC_1$.
%
The child supertag, $tag_c$, is embedded using the target supertag embedding
$G_s$ and concatenated with the lexical vector, $\omega_{lex}$, forming an intermediate vector representation of the node, $\omega_{node}$.
%
The node vector is repeated for each of the parent's possible supertags, $tagset_p$, and then concatenated with their embeddings to construct the set of treelet vectors, $\Omega_{treelet}$.
%
The vector states for the leaf nodes are similarly constructed, but instead combine the lexical vector, $\omega_{lex}$ with the embeddings of the child's possible supertags, $tagset_c$.
%
The final operation induces a probability distribution over the treelet and leaf vectors using a score computed by the vectorized function, $\Psi_{predict}$, as the scalar in a softmax distribution. 
%
%\vspace*{-0.55cm}
\begin{align}
&\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
&\omega_{node}=concat([G_s(tag_c);~\omega_{lex}]) \nonumber \\
&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_s(tagset_p)]) \nonumber \\
&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_s(tagset_c)]) \nonumber \\
&P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
\frac{exp(\Psi_{predict}(\omega_{treelet_i})))}
{\sum_{j \in |tagset_p|} exp(\Psi_{predict}(\omega_{treelet_j})))} \nonumber \\
&P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
\frac{exp(\Psi_{predict}(\omega_{leaf_i})))}
{\sum_{j \in |tagset_c|} exp(\Psi_{predict}(\omega_{leaf_j})))} \nonumber 
\end{align}
%\vspace*{-0.45cm}

At generation time, we are given a full dependency tree.  A decoding
step is necessary to compute a high probability assignment for all
supertags simultaneously.
%
In this process, tags for children must be chosen consistently with
one another, and the resulting probabilistic information must be
propagated upward to rerank tags elsewhere in the tree.
%
We solve this problem with an A* algorithm.
%
At each step, the algorithm uses a priority queue to select subtrees based on their inside-outside scores.
%
The inside score is computed as the sum of the log probabilities of the supertags in the subtree.
%
The outside score is the sum of the best supertag for nodes outside the subtree, similar to \newcite{lewis2014improved}.
%
Once selected, the subtree is attached to the possible supertags of its parent that are both locally consistent and consistent among its already attached children.
%
These resulting subtrees are placed into the priority queue and the algorithm iterates to progress the search.
%
The search succeeds when the first complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed.}

@@

The Fergus-R model uses only lexical information as input to calculate the probability distribution over each node's supertags. 
%
The specific formulation is detailed in Eq. \ref{eq:fergusr}.
%
First, a parent node's children, $lex_C$, are embedded using the word embedding matrix, $G_w$, and then summarized with an attention function, $\Psi_{attn}$, to form the child context vector, $\omega_{C}$. 
%
The child context is concatenated with the embedded lexical information of the parent node, $lex_p$, and mapped to a new vector space with a fully connected layer, $FC_1$, to form the lexical context vector, $\omega_{lex}$.
%
The context vector and a topological vector for indexing the internal tree state (see \ref{subsec:rtn}) are passed to the recurrent tree network, $RTN$, to compute the full state vector for the parent node, $\omega_{node}$.
%
Similar to Fergus-N, the state vector is repeated and concatenated with the vectors of the parent node's possible supertags, ${tagset}_p$, and mapped to a new vector space with a fully connected layer, $FC_2$.
%
A vector in this vector space is labeled $\omega_{elementary}$ because the combination of supertag and lexical item constitutes an elementary tree.
%
The last step is to compute the probability of each supertag using the vectorized function, $\Psi_{predict}$.
%
%\vspace*{-0.55cm}
\begin{align}
&\omega_{C} = \Psi_{attn}(G_w(lex_C)) \label{eq:fergusr} \\
&\omega_{lex} = FC_1(concat(\omega_{C},~G_w(lex_p))) \nonumber \\
&\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_s(tagset_p))) \nonumber \\
&P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
\frac{exp(\Psi_{predict}(\omega_{elementary_i})))}
{\sum_{j \in |\Omega|} exp(\Psi_{predict}(\omega_{elementary_j))}} \nonumber
\end{align}
%\vspace*{-0.45cm}

@@

The final step to linearizing the output of Fergus-N and Fergus-R---a dependency tree annotated with supertags and partial attachment information---is a search over possible orderings with a language model. 
%
There are many possibilities, primarily due to ambiguities in insertion
order.
%
Following \newcite{bangalore2000exploiting}, a language model is used to select between the alternate orderings. 
%
The language model used is a two-layer LSTM trained using the Keras library on
the surface form of the Penn Treebank.
%
The surface form was minimally cleaned\footnote{With respect to the surface form, the only
cleaning operations were to merge proper noun phrases into single tokens.  Punctuation and other
common cleaning operations were not performed.} to simulate realistic scenarios.

@@

Using the representations of \ref{sec:neural}, the models of
\ref{sec:models} can be instantiated in six different ways.  We can
use a feed-forward Fergus-N architecture or a recurrent Fergus-R
architecture.  Each architecture can embed supertags \emph{minimally},
by learning a scalar corresponding to each supertag; \emph{atomically}, by learning an
embedding vector corresponding to each supertag; or
\emph{structurally}, by using convolutional coding over each
supertag's tree structure to form a vector.
%
In each case, the vector (a size-one vector in the minimal condition) is concatenated as described in \ref{sec:models}.

@@

From this input, we extracted the atomic supertag prediction instances
and trained a network defined by each of the architectures of
\ref{sec:models} and each of the supertag representations of
\ref{sec:neural}.  As always, we used Sections 02-21 for training,
Section 22 for development, and Section 23 for testing.  A complete
description of network organization and training parameters is given
in the appendix.  The code and complete experimental setup are
publicly
available.\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}

@@

Figure~\ref{table:linresults} shows the NLG evaluation results for the
different models.
%
All differences in model are significant using an
Independent\footnote{An Independent t-test was used instead of a
  Paired-Sample t-test because of intermittent failures during
  linearization that resulted in slightly different numbers of
  observations.} t-test ($p<10^{-5}$).
%
For both models, the differences between structural embeddings (using convolutional coding) and atomic embeddings (using standard vector embedding techniques) were not significant, while the differences between the two embeddings and minimal embeddings were significant ($p<10^{-5}$). 
%
The performance confirms our expectation that differences in supertag
accuracy after decoding correlate with NLG accuracy overall, but that
differences in NLG performance are attenuated.
%
We note by comparison that Bangalore and Rambow report an accuracy of
74.9\% in their best evaluation of FERGUS---on a data set of just 100
sentences with an average length of 16.7.
%
Our evaluation, on 2400 sentences with an average length of 22, is
more strenuous.

@@

There are several lines of related work which explore stochastic tree
models from the standpoint of parsing and understanding.
%
While using the same methods, NLG has different goals and we think the perspective is instructive. 
%
Where parsing infers the most probable underlying structure,
generation infers the most likely way of expressing a semantic
structure.
%
This divergence of goals leads to different concerns, alternatives, and emphasis. 
%---which unlabeled dependency trees have served as our proxy in this work.

The works most similar to ours explicitly model tree structures, but focus on resolving the uncertainty involved with the latent structure of an observed sentence. 
%
For example, the top down tree structure of \newcite{zhang2016top} expresses the generation of a dependency structure as the decisions of a set of long short-term memory networks.  
%
For each decision, the possible options are different tree structures which can produce the target linear form.
%
In contrast, the generation problem is concerned with different linear forms that can result from the same tree structure. 
%
In more extensive tasks, the generation problem can include simulated interpretation to inform decisions; using the ease of structural inference from linear form quantifies the understandability of a sentence.
